tl22-01_analysingNetworkStats

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingNetworkStats.Rmd
Add many required vivax files and change paths in vivax scripts

```{r read_data}
accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
assembly.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
subfam.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-02_VirSubfamNetwork/pir_info/lopez13_vir_subfam.txt', header = FALSE)
colnames(subfam.df) <- c('Geneid','subfam')
expt.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/scripts/all-cat.list', header = FALSE, sep = '')
colnames(expt.df) <- c('Accession', 'expt')
expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')
```

```{r MCL_parameters}
# list.files('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
  #          pattern = 'mci.+list') %>% 
  # head %>% 
  # str_extract_all(pattern = '(?<=_)[[:digit:]]+(?=pid)|[[:digit:]]+(?=lngth)|(?<=mci\\.)I.+(?=\\.list)')

network_file_dir <- '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1'

@@

```{r get_MCLclusters}
fasta <- read.fasta(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

@@

# 
# write.table(sequence.df,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')

sequence.df <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt',
                          check.names = FALSE)

@@

# pid_lntgh.df <- lapply(perc_ident.vec, function(perc_ident){
#   lapply(length_aa.vec, function(length_aa){
#     #Read in BLAST file for these percentage id and length aa 
#     # Get a list of the unique sequences from the first two columns
#     blast.names <- read.delim(paste0(
#       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
#       perc_ident,'pid',
#       length_aa,'lngthNoSame2Same.blast'
#     ),
#     header = FALSE) %>% 
#       .[,c(1,2)] %>% 
#       unlist() %>% 
#       unique
#     #lapply across the sequences in the fasta to record whether they are present in this network (i.e. in the BLAST file)
#     in_network.df <- lapply(names(fasta), 
#            function(seq.id){
#              in_network <- ifelse(seq.id %in% blast.names,
#                                   1,
#                                   0)
#              return(data.frame(label = seq.id, 
#                                in_network = in_network,
#                                pid = perc_ident,
#                                lngth = length_aa))
#            }) %>% bind_rows
#   })
# }) %>% bind_rows

@@

# write.table(sequence.df.melt,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
sequence.df.melt <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
```

@@

# node.df <- filter(sequence.df.melt,
#        pid == 30,
#        lngth == c(25)) %>% 
#   mutate(network_parameters = paste0(pid,'pid',lngth,'lngthI',inflation)) %>% 
#   dcast(label + Id + source + subfam + expt ~ network_parameters, value.var = 'cluster_num')
# 
# write.csv(node.df,
#           '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
#           row.names = FALSE,
#           quote = FALSE)

node.df <- read.csv(
  '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')

@@

# write.csv(cluster_subfam.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
#           row.names = FALSE,
#           quote = FALSE)

@@

# blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
#            header = FALSE)
# 
# node_labels.df <- blast_30pid25lngth.df %>%
#   rowwise() %>%
#   mutate(label1 = node.df$Id[node.df$label %in% V1],
#          label2 = node.df$Id[node.df$label %in% V2]) %>%
#   select(V1,V2,label1, label2)
# 
# node_labels_noDup.df <- node_labels.df[!duplicated(apply(node_labels.df,
#                                                          sort,
#                                                          MARGIN =1),
#                                                    MARGIN=2),]
# #Already appears to have same2same matches absent.
# 
# node_labels_noDup_toSave.df <- select(node_labels_noDup.df, label1, label2)
# colnames(node_labels_noDup_toSave.df) <- c('Source', 'Target')

# write.csv(node_labels_noDup_toSave.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
#           row.names = FALSE,
#           quote = FALSE)

@@

                  if(global_save_var){
                    ggsave(plot = plot,
                           paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptComposition_',
                                  include,'_',mcl_name,'.pdf'),
                           device = 'pdf',
                           width = 7,
                           height = 5,
                           units = "in")
                  }

@@

if(global_save_var){
  ggsave(plot = plot,
         paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_clustComp_Expt.pdf'),
         device = 'pdf',
         width = 7,
         height = 8,
         units = "in")
}
```

@@

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptStep_'
                         ,mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

@@

lapply(unique(node.melt$mcl), 
       function(mcl_name){
         mcl_num_subfamNum.df <- filter(node.melt, 
                                        mcl == paste(mcl_name)) %>% 
           #Make clust_num numeric for correct ordering by number.
           mutate(clust_num = as.numeric(clust_num),
                  subfam = ifelse(subfam == 'Sal1', 
                                  'unassigned',
                                  subfam)) %>% 
           group_by(subfam, clust_num) %>% 
           filter(clust_num %in% 1:25, 
                  source == 'Sal1') %>%
           dplyr::summarize(num = n_distinct(label))

@@

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_sal1Composition_',
                         mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

@@

```{r pir_sizes}

node_length.melt <- filter(node.melt, 
       mcl == 'X30pid25lngthI1.4') %>% 
  rowwise() %>% 
  mutate(orf_length = length(getElement(fasta, label)))

@@

if(global_save_var){
  ggsave(plot = plot_orf,
         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_ORFlength.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

@@

Cluster 8 particularly long?

Looks like there's a little group of longer pirs for some experiments/P01/Sal1.
Like over 800aa - pretty much all in Cluster 1 (E). 
I wonder if this is the two groups in the Gephi network.

Cluster 7 and 9 low variability, potentially shorter?

```{r kwtest_length}
#K-W test of if there is a signif difference between the clusters for length

kruskal.test( node_length.melt$orf_length, factor(node_length.melt$clust_num))
kruskal.test(node_length_clust1.11_noZhu16.melt$orf_length,
             factor(node_length_clust1.11_noZhu16.melt$clust_num))

clust_comp <- combn(1:11, m = 2, simplify = FALSE)

#Do the tests for each combination, combione into a data frame and use p.adjust multiple testing adjustment.
( length_tests.df <- lapply(c('greater', 'less'), 
                            function(wilcox_text_alt){
                              lapply(clust_comp, function(comparison){
  test.stat <- wilcox.test(orf_length ~ clust_num,
                           data = filter(node_length.melt,
                                         expt != 'Zhu16_known'),
                           subset = clust_num %in% comparison,
                           alternative = wilcox_text_alt)
  return(data.frame(comparison_1 = comparison[1],
                    comparison_2 = comparison[2],
                     stat = test.stat$statistic,
                     pval = test.stat$p.value,
                    wilcox_text.alt = wilcox_text_alt))
                              }) %>% bind_rows
}) %>% 
  bind_rows %>% 
  mutate(padj = round(p.adjust(pval), digits = 5),
         pval = round(pval, digits = 5)) %>% 
  arrange(desc(padj)) )

length_med.df <- node_length.melt %>% 
  group_by(clust_num) %>% 
  summarise(orf_length_med = median(orf_length))

length_tests.mat <- mutate(length_tests.df,
                           orf_ratio = ifelse(padj <= 0.05, 
                                              length_med.df$orf_length_med[
                                                match(comparison_1, 
                                                      length_med.df$clust_num)
                                              ]/length_med.df$orf_length_med[
                                                match(comparison_2, 
                                                      length_med.df$clust_num)
                                              ], 
                                              NA)) %>% 
  dcast(.,
        comparison_1 ~ comparison_2,
        value.var = 'orf_ratio')
rownames(length_tests.mat) <- length_tests.mat[,1]
length_tests.mat <- length_tests.mat[,-1]

# nams <- unique(unlist(clust_comp))
# 
# #Need to add the first and last rows due to combn not using 1 as combination_2 and 11 as combination_1.
# one.mat <- matrix(nrow = nrow(length_tests.mat),
#        dimnames = list(rep(NA, length.out = nrow(length_tests.mat)), first(nams))) 
# length_tests.mat <- cbind(one.mat, length_tests.mat)
# 
# last.mat <- matrix(ncol = ncol(length_tests.mat),
#        dimnames = list(last(nams), rep(NA, 
#                                        length.out = ncol(length_tests.mat))))
# colnames(last.mat) <- names(length_tests.mat)
# length_tests.mat <- rbind(length_tests.mat, last.mat)
# 
# library(ComplexHeatmap)
# library(viridis)
# #Make matrix symmetrical
# length_tests.mat[
#   lower.tri(length_tests.mat)
# ] <- 1/(t(length_tests.mat)[
#   lower.tri(length_tests.mat)
# ])
# 
# colnames(length_tests.mat) <- rownames(length_tests.mat) <- nams
# 
# Heatmap(length_tests.mat, 
#         cluster_columns = FALSE, 
#         cluster_rows = FALSE,
#         na_col = 'black',
#         col = circlize::colorRamp2(c(0, 1, 1.4), c('blue', 'white', 'red')))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '^1$|3|8|10'))) %>%
  arrange(desc(padj))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '7'))) %>%
  arrange(desc(padj))
length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '9'))) %>%
  arrange(desc(padj))
```

@@

  #Read in BLAST file 
blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
           col.names = outfmt6_headers) %>% 
  mutate(
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  )

@@

```{r}
vir14_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVX_108770') | sseqid %in% 
           c('PVX_108770'))
hist(vir14_blast$pident, breaks = 100)
hist(vir14_blast$scovhsp, breaks = 100)
hist(vir14_blast$qcovhsp, breaks = 100)

>>>>

```{r read_data}
accession_table <- read.csv('samples_to_cat.csv')
assembly.df <- read.delim('tl22-01_assembly_info.txt')
subfam.df <- read.delim('lopez13_vir_subfam.txt', header = FALSE)
colnames(subfam.df) <- c('Geneid','subfam')
expt.df <- read.delim('all-cat.list', header = FALSE, sep = '')
colnames(expt.df) <- c('Accession', 'expt')
expt_num.df <- read.csv('expt_codes.csv')
```

```{r MCL_parameters}
# list.files('hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
  #          pattern = 'mci.+list') %>% 
  # head %>% 
  # str_extract_all(pattern = '(?<=_)[[:digit:]]+(?=pid)|[[:digit:]]+(?=lngth)|(?<=mci\\.)I.+(?=\\.list)')

network_file_dir <- 'networks'

@@

```{r get_MCLclusters}
fasta <- read.fasta(file = 'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

@@

# 
# write.table(sequence.df,
#             file = 'networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')

sequence.df <- read.table(file = 'sequence-MCLclusters_tpmOver1.txt',
                          check.names = FALSE)

@@

# pid_lntgh.df <- lapply(perc_ident.vec, function(perc_ident){
#   lapply(length_aa.vec, function(length_aa){
#     #Read in BLAST file for these percentage id and length aa 
#     # Get a list of the unique sequences from the first two columns
#     blast.names <- read.delim(paste0(
#       'hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
#       perc_ident,'pid',
#       length_aa,'lngthNoSame2Same.blast'
#     ),
#     header = FALSE) %>% 
#       .[,c(1,2)] %>% 
#       unlist() %>% 
#       unique
#     #lapply across the sequences in the fasta to record whether they are present in this network (i.e. in the BLAST file)
#     in_network.df <- lapply(names(fasta), 
#            function(seq.id){
#              in_network <- ifelse(seq.id %in% blast.names,
#                                   1,
#                                   0)
#              return(data.frame(label = seq.id, 
#                                in_network = in_network,
#                                pid = perc_ident,
#                                lngth = length_aa))
#            }) %>% bind_rows
#   })
# }) %>% bind_rows

@@

# write.table(sequence.df.melt,
#             file = 'networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
sequence.df.melt <- read.table(file = 'sequence-MCLclusters_tpmOver1_longForm.txt')
```

@@

# node.df <- filter(sequence.df.melt,
#        pid == 30,
#        lngth == c(25)) %>% 
#   mutate(network_parameters = paste0(pid,'pid',lngth,'lngthI',inflation)) %>% 
#   dcast(label + Id + source + subfam + expt ~ network_parameters, value.var = 'cluster_num')
# 
# write.csv(node.df,
#           'networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
#           row.names = FALSE,
#           quote = FALSE)

node.df <- read.csv(
  'sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')

@@

# write.csv(cluster_subfam.df,
#           file = 'networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
#           row.names = FALSE,
#           quote = FALSE)

@@

# blast_30pid25lngth.df <- read.delim('hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
#            header = FALSE)
# 
# node_labels.df <- blast_30pid25lngth.df %>%
#   rowwise() %>%
#   mutate(label1 = node.df$Id[node.df$label %in% V1],
#          label2 = node.df$Id[node.df$label %in% V2]) %>%
#   select(V1,V2,label1, label2)
# 
# node_labels_noDup.df <- node_labels.df[!duplicated(apply(node_labels.df,
#                                                          sort,
#                                                          MARGIN =1),
#                                                    MARGIN=2),]
# #Already appears to have same2same matches absent.
# 
# node_labels_noDup_toSave.df <- select(node_labels_noDup.df, label1, label2)
# colnames(node_labels_noDup_toSave.df) <- c('Source', 'Target')

# write.csv(node_labels_noDup_toSave.df,
#           file = 'networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
#           row.names = FALSE,
#           quote = FALSE)

@@

                  if(global_save_var){
                    ggsave(plot = plot,
                           paste0('plots/networks_exptComposition_',
                                  include,'_',mcl_name,'.pdf'),
                           device = 'pdf',
                           width = 7,
                           height = 5,
                           units = "in")
                  }

@@

if(global_save_var){
  ggsave(plot = plot,
         paste0('plots/networks_clustComp_Expt.pdf'),
         device = 'pdf',
         width = 7,
         height = 8,
         units = "in")
}
```

@@

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('plots/networks_exptStep_'
                         ,mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

@@

lapply(unique(node.melt$mcl), 
       function(mcl_name){
         mcl_num_subfamNum.df <- filter(node.melt, 
                                        mcl == paste(mcl_name)) %>% 
           #Make clust_num numeric for correct ordering by number.
           mutate(clust_num = as.numeric(clust_num),
                  subfam = ifelse(subfam.x == 'Sal1', 
                                  'unassigned',
                                  subfam.x)) %>% 
           group_by(subfam, clust_num) %>% 
           filter(clust_num %in% 1:25, 
                  source == 'Sal1') %>%
           dplyr::summarize(num = n_distinct(label))

@@

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('plots/networks_sal1Composition_',
                         mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

@@

```{r pir_sizes}
node_length.melt <- filter(node.melt, 
       mcl == 'X30pid25lngthI1.4') %>% 
  rowwise() %>% 
  mutate(orf_length = length(getElement(fasta, label)))

@@

if(global_save_var){
  ggsave(plot = plot_orf,
         'plots/networks_ORFlength.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

@@

@@

  #Read in BLAST file 
blast_30pid25lngth.df <- read.delim('all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
           col.names = outfmt6_headers) %>% 
  mutate(
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  )

@@

```{r vir14}
vir14_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVX_108770') | sseqid %in% 
           c('PVX_108770'))
hist(vir14_blast$pident, breaks = 100)
hist(vir14_blast$scovhsp, breaks = 100)
hist(vir14_blast$qcovhsp, breaks = 100)
